---
output: github_document
editor_options: 
chunk_output_type: console
---
## package development
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) ![R](https://img.shields.io/badge/R-%3E%3D%202.10-blue) 
![R CMD check](https://img.shields.io/badge/R%20CMD%20check-passing-green)
![Coverage](https://img.shields.io/badge/coverage-20%25-red)

## morphalr: Morphological Analysis for Archaeology
A package to compute morphological indices of spatial entities (e.g. parcels, buildings). It also provides visuals of multivariate statistics results (ACP and HCA).


## Installing
Package currently exist as development on github.

Install package from github:
```{r, eval=FALSE}
library(remotes)
install_github(repo = "JGravier/morphalr")
```

## Compute orientations from polygon
Load data of parcels of the city of Rouen in 1827 and extract segments (lines) from polygons.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(morphalr)

# transform polygons as segments (lines)
linerouen <- morphalr_geom_to_segment(sfobject = rouen_1827, to = 'LINESTRING')
```
`morphal_geom_to_segment()` can be applied on LINESTRING, e.g. in the case of streets modeled as lines.

Compute orientations of segments with East looking and perpendicular parameters.
```{r, message=FALSE, warning=FALSE}
orientationsest <- morphalr_segment_orientation(sfsegments = linerouen, looking = 'E', perpendicular = TRUE)

orientationsest |>
  ggplot() +
  geom_sf(aes(color = orientation)) +
  scale_color_viridis_c() +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

## Compute morphological indices
Example with distance to minimal bounding rectangle (DSR):
```{r}
rouen_1827 |>
  morphalr_dsr() |>
  ggplot() +
  geom_sf(aes(fill = dsr), color = 'grey90', linewidth = 0.05) +
  scale_fill_viridis_c(direction = -1) +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

