---
title: "morphalr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{morphalr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(morphalr)
library(ggplot2)
```

## Compute orientations
Load data of parcels of the city of Rouen in 1827 and extract segments (lines) from polygons.
```{r, message=FALSE, warning=FALSE}
# transform polygons as segments (lines)
linerouen <- morphalr_geom_to_segment(sfobject = rouen_1827, to = 'LINESTRING')
```
`morphal_geom_to_segment()` can be applied on LINESTRING, e.g. in the case of streets modeled as lines.

Compute orientations of segments with East looking and perpendicular parameters.
```{r, message=FALSE, warning=FALSE}
orientationsest <- morphalr_segment_orientation(sfsegments = linerouen, looking = 'E', perpendicular = TRUE)

orientationsest |>
  ggplot() +
  geom_sf(aes(color = orientation)) +
  scale_color_viridis_c() +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), panel.grid = element_blank())
```

## Compute morphological indices of polygons
Example with distance to minimal bounding rectangle (DSR):
```{r, message=FALSE, warning=FALSE}
rouen_1827 |>
  morphalr_dsr() |>
  ggplot() +
  geom_sf(aes(fill = dsr), color = 'grey90', linewidth = 0.05) +
  scale_fill_viridis_c(direction = -1) +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), panel.grid = element_blank())
```

## Compute clusters from indices
Compute indices.
```{r, message=FALSE, warning=FALSE}
rouen_with_indices <- rouen_1827 |>
  morphalr_circularity() |>
  morphalr_dsc() |>
  morphalr_dsr()

rouen_with_indices <- rouen_with_indices[,2:5] # without rowid
rouen_with_indices
```

Create HCA clustering and plot result:
```{r, message=FALSE, warning=FALSE}
clusterrouen <- morphalr_clustering(sf = rouen_with_indices, pca_center = TRUE, pca_scale = TRUE, hca_method = 'ward.D2')

plot(clusterrouen, labels = FALSE, hang = 0)
```

### Clusters
Cartography of clusters:
```{r, message=FALSE, warning=FALSE}
morphalr_clusters(sf = rouen_with_indices, clustering = clusterrouen, cutting = 6) |>
  ggplot() +
  geom_sf(aes(fill = cluster), color = 'white', linewidth = 0.02) +
  ggthemes::scale_fill_colorblind() +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), panel.grid = element_blank())
```

Summary of clusters with mean of center-scale values of variables by cluster.
```{r, message=FALSE, warning=FALSE}
morphalr_clusters_mean(sf = rouen_with_indices, clustering = clusterrouen, cutting = 6)
```
