---
title: "Computing orientations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing orientations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, attr.warning=FALSE, message=FALSE}
library(mafa)
library(parallel)
library(data.table)
library(dplyr)
library(sf)
library(ggplot2)
```

## Extracting segments from sf
Extracting segments with `morphal_geom_to_segment()` can be applied to polygon or linestring geometries. Default output consists of the input sf composed of segments (i.e LINE as line composed of only two points) in a MULTILINESTRING geometry.
```{r, message=FALSE, warning=FALSE}
# transform polygons as segments (lines)
mafa_geom_to_segment(sfobject = rouen_1827)
```

Parameter `to` can provides LINESTRING output.
```{r, message=FALSE, warning=FALSE}
rouenline <- mafa_geom_to_segment(sfobject = rouen_1827, to = 'LINESTRING')
rouenline
```


## Parallelizing extraction
Computing of `morphal_geom_to_segment()` can be long and it may be useful to apply parallel computing.
```{r, message=FALSE, warning=FALSE, eval=FALSE}
nbcores <- parallel::detectCores() - 1
rouensplit <- rouen_1827 |>
  mutate(bloc1000 = 1:nrow(rouen_1827) %/% 1000) |>
  group_by(bloc1000) |>
  group_split()

rouenparall <- mclapply(X = rouensplit, FUN = mafa_geom_to_segment, mc.cores = nbcores)
rouenparall <- data.table::rbindlist(rouenparall) |>
  st_as_sf() |>
  select(-bloc1000)

# as LINESTRING
rouenparall <- rouenparall |>
  st_cast(to = 'LINESTRING')
```

## Computing orientations
Compute orientations of segments with `mafa_segment_orientation()`. The looking default parameters is "N" for North, and results are in degrees on a semicircle [-90;90].
```{r, message=FALSE, warning=FALSE}
orientations <- mafa_segment_orientation(sfsegments = rouenline)

orientations |>
  ggplot() +
  geom_sf(aes(color = orientation)) +
  scale_color_viridis_c() +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), panel.grid = element_blank())
```


Compute orientations of segments with East looking and perpendicular parameters. The latter makes it possible to combine perpendicular orientations, with values in degrees recalculated in a quarter circle [0;90].
```{r, message=FALSE, warning=FALSE}
orientations <- mafa_segment_orientation(sfsegments = rouenline, looking = 'E', perpendicular = TRUE)

orientations |>
  ggplot() +
  geom_sf(aes(color = orientation)) +
  scale_color_viridis_c() +
  theme_bw() +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), panel.grid = element_blank())
```

